---
title: "Choosing the right visualization"
subtitle: "ANLY503 Advanced Data Visualization"
institution: "Georgetown University"
author: "Marck Vaisman & Abhijit Dasgupta"
date: "Spring 2022"

---

```{r child="../format/setup-options.Rmd"}
```

```{r, include=FALSE}
currdir <- here('slides','yau2-3')
imgdir <- path(currdir,'img')

options(htmltools.dir.version = FALSE, htmltools.preserve.raw = FALSE) # required for JS graphics

knitr::opts_chunk$set(cache=TRUE)
```


## Last week

Designing for an audience vs. designing for you

.pull-left[
**Tufte**
1. Show data variation, not design variation
1. Do not use graphics to quote data out of context
1. Use clear, detailed, thorough labelling.
1. Representation of numbers should be directly proportional to numerical quantities
1. Don't use more dimensions than the data require
]

.pull-right[
**Nathan Yau**
1. Check the data
1. Explain encodings
1. Label axes
1. Include units
1. Keep your geometry in check
1. Include your sources
1. Consider your audience
]


---

## Today

1. Choosing the right visualization
  * All about asking questions
  * Decomposing your chart
  * Understanding encodings
1. Visualization critique discussion
1. Final conceptual and design considerations
  * Making readable graphics
  * Deep dive into Tufte's principles
  * Bringing it all together. 
1. **Case study 2** presentation
1. Activity: Exploring Python's static graphing packages
  * matplotlib
  * seaborn



---

```{r child='selecting-a-chart.Rmd'}
```

---

```{r child='breaking-it-up.Rmd'}
```

---

```{r child='visual-encodings.Rmd'}
```

---

```{r child='what-is-wrong.Rmd'}
```

---

```{r child='layers.Rmd'}

```

---

```{r child='tufte.Rmd'}
```

---

```{r child='bringing-it-together.Rmd'}
```


